\textsf{Are frontier models superhuman chemists?}

### 2024-03-19 src/tex/ms.tex
write reading ease to file

In \Cref{fig:confidence_vs_performance} we show that for some models there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\autocite{Li_2023}
For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of XX (on a scale of 1--5) for the XX (out of XX) questions it answered in correctly (XX for the XX questions it answered correctly).

@@

To maintain the consistency with the training, we employ distinct prompt templates in general tailored for completion models and instruction-tuned models. 
We pose instruction or impose constraints on the models within these templates to receive response in a specific format so that a robust, fair and consistent parsing can be performed as explained in \Cref{sec:parsing}.
Certain models are trained with special annotations and LaTeX syntax for scientific notations, chemical reactions, or symbols embedded within the text. 
For example, all the SMILES representations are encapsulated within \texttt{[START\_SMILES][\text backslash END\_SMILES]} in Galactica\autocite{taylor2022galactica}.
Our prompting strategy consistently adheres to  these details in a model-specific manner by post-processing \LaTeX syntax, chemical symbols, chemical equations and physical units (by either adding or removing wrappers).

\paragraph{Parsing}
Our parsing workflow is, by default, multistep and primarily based on regular expressions.
In the case of instruction-tuned models we first attempt to identify the \texttt{[ANSWER][\textbackslash ANSWER]} environment we prompt the model to report the answer in.
In the case of completion models, this step is skipped. From there, we attempt to extract the relevant enumeration letters (for multiple-choice questions) or numbers.
In the case of numbers, our regular expression was engineered to be able to deal with various forms of scientific notation.
As initial tests indicated that models sometimes return integers in the form of words, e.g. \enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion.
In case these hard-coded parsing steps fail, we fall back to using a \gls{llm}, e.g. Claude-2, to parse the completion.
Manual verification indicates that this \gls{llm}-based parsing is not a relevant error source and has been performed correctly in all cases we checked (randomly sampled 4 questions per topic).

>>>>

In \Cref{fig:confidence_vs_performance} we show that for some models there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
For applications in which humans might rely on the models to provide answers, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the models' outputs.\autocite{Li_2023}
For example, for the questions about the safety profile of compounds, GPT-4 reported an average confidence of \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_correct_overall.txt} questions it answered in correctly and \variable{output/model_confidence_performance/gpt4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading for the questions about \gls{ghs} pictograms with an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.

@@

To maintain the consistency with the training, we employ distinct prompt templates in general tailored for completion models and instruction-tuned models. 
We pose instruction or impose constraints on the models within these templates to receive response in a specific format so that a robust, fair and consistent parsing can be performed as explained below.
Certain models are trained with special annotations and LaTeX syntax for scientific notations, chemical reactions, or symbols embedded within the text. 
For example, all the SMILES representations are encapsulated within \texttt{[START\_SMILES][\text backslash END\_SMILES]} in Galactica\autocite{taylor2022galactica}.
Our prompting strategy consistently adheres to  these details in a model-specific manner by post-processing \LaTeX syntax, chemical symbols, chemical equations and physical units (by either adding or removing wrappers).

\paragraph{Parsing}
Our parsing workflow is, by default, multistep and primarily based on regular expressions.
In the case of instruction-tuned models we first attempt to identify the \texttt{[ANSWER][\textbackslash ANSWER]} environment we prompt the model to report the answer in.
In the case of completion models, this step is skipped. From there, we attempt to extract the relevant enumeration letters (for multiple-choice questions) or numbers.
In the case of numbers, our regular expression was engineered to be able to deal with various forms of scientific notation.
As initial tests indicated that models sometimes return integers in the form of words, e.g. \enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion using regular expressions.
In case these hard-coded parsing steps fail, we fall back to using a \gls{llm}, e.g. Claude-2, to parse the completion.
Manual verification indicates that this \gls{llm}-based parsing is not a relevant error source and has been performed correctly in all cases we checked (randomly sampled 4 questions per topic).
